PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2023-03-31 refereeResponse.txt

418: I would specifically state that for the large alpha cases runaway growth only occurs at the start, to avoid initial confusion that the slope turns around when the planetesimal supply is depleted.
435: Is this true? I was under the impression that the residual planetesimal population remains? I think this does not seem consistent with the rest of the work...?
	TODO: Change the wording here. Maybe this isnt true. Would the growth mode simply 'flatten' out here?
446: What exactly does alpha being non-negligible mean? Given that alpha is always less than unity, what alpha value marks the transition of the oligarchic growth regime vs, consumption regime? See major comment.
	This is a good point, and I'm not sure if we have a definitive answer. Inspecting figure 11, theh transition seems to occur near alpha ~ 0.1. I think the best thing to do here is to state the converse: for alpha << 1, most encounters result in scattering and you get oligarchic growth. The key finding from this section is that oligarchic growth does not occur when alpha is not << 1. The text has been updated to reflect this.

5. Full disk Simulation
5.1 Initial Conditions
470: Reference of TRAPPIST-1
	Added reference to Gillon 2016 + 2017, Agol 2021
483 I recommend writing Eq. (6) along the lines of 10 g/cm^2 \times A (M_*/M_\odot)(r/1AU)^(-delta) for readability. The Sigma_MMSN is never used so it's not necessary.
	Done
Table 1.
I don't think the notes are necessary
I would probably put the planetesimal masses in units of Ceres
	Where should I put the units?
I think Table 1 and Table 2 could be combined
	Table 1 has been merged into table 2.

490: How could such an enhancement in planetesimal surface density be achieved initially? Is this initial condition consistent with planetesimal formation models, e.g. Dra̧żkowska & Dullemond 2018, Lenz et al 2019.
	This is briefly discussed in the introduction but we will expand upon it here. Most mechanisms involve dust traps or pressure bumps occuring at some preferential location in the disk which trigger the streaming instability. Dra̧żkowska & Dullemond 2018 show that a migrating snow line can actually trigger planetesimal formation over a fairly wide area, and also produce surface density enhancements at least as large as the ones we use. Given that low-mass stars have particularly extreme luminosity changes during the pre-main sequence (see Baraffe 2015), one can imagine that it might not be hard to trigger planetesimal formation over a wide area. This particular paragraph in the text has been expanded to discuss this.

493: What is the motivation behind variation in the power-law index? Is there significant uncertainty in the MMSN power law index?
	Chiang and Laughlin 2013 applied a similar analysis to Kepler multis and found that delta varies between 1.6 and 1.8. Sub-mm observations of cold outer disks finds power law slopes as low as 0.5 (Mundy 2000, Andrews 2009, 2010). Therefore, we vary delta by +/- 1.0 from the MMSN value. There is now a separate paragraph in the text explaining this
499: Maybe elaborate on why this equilibrium is physical? Are there other works which use the same balance to inform initial conditions?
	t_vs = t_gas is more of an upper limit on the eccentricity, although this is the same technique that Kokubo + Ida 1998 use to set up their initial conditions. It's certainly possible that the disk could have started out dynamically colder, although as we showed in figure 2, this doesn't seem to affect the resulting protoplanets. We have added a short discussion of these points to the text.
504: missing respectively. And semi-major axes.
	Fixed
505: When is the Stokes regime valid, and what does it mean?
	In the Stokes regime, the mean free path of the gas is much smaller than the size of the bodies and the gas can be treated as a fluid. For conditions in a typical protoplanetary disk, Stokes drag is relevant for cm-sized bodies and larger. This has been added to the text.
506: The letter tau is used for a timescale just above, but now it says t. I would recommend staying consistent. Tau tends to be used for dimensionless timescales, so I would probably go with t_s and t_vs.
	tau_vs has been changed to t_vs in equation 7.
511: Solid-to-gas surface density ratio of 1/240 seems very small, especially given that the authors are considering the solid surface density to already be enhanced by a factor of 100. Also, if the planetesimals are already formed, it is conceivable that the gas disk is already dissipating (especially so close in). I think the authors could be over-estimating the gas density by up to a factor of a 100, which means the stopping time would be a factor of 100 longer.
	TODO: Fix
516: I suggest using another symbol for the temperature power law, since Q is associated with the Toomre value of the disk, for example q. Also, this equation should read T(r) = T_0(r/1AU)^-q, right?
	Q was changed to q to be consistent with the Andrews + Williams paper. the '/1AU' has been added to the T(r) equation.
521: Where is that equation coming from? Maybe add reference.
	This is equation 4.30 from Phil Armitage's 'Astrophysics of Planet Formation'. A reference has been added to the text.

### 2023-03-31 main.tex
Referee edits up to sec 5.2

where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g
cm$^{-2}$ is the surface density of the minimum-mass solar nebula
\citep{hayashi81} at 1 AU, $A$ is an enhancement factor and $\delta$ is the power law index.
In the first case (fdHi), we model a disk that follows a MMSN power law slope ($\delta$ = 1.5), with the overall normalization enhanced by a factor of 100. This choice of normalization for the solid surface density profile appears necessary in order to reproduce many observed short period terrestrial worlds in-situ \citep{hansen12}. Additionally, we vary the power law index (fdHiShallow, fdHiSteep) and overall normalization (fdLo) of $\Sigma(r)$.

@@

\begin{equation}\label{eq:vs_timescale}
    \tau_{vs}  = \frac{\left< e^2 \right>}{d \left< e^2 \right> / dt} \approx \frac{1}{40}\left(\frac{\Omega^{2} a^{3}}{2 G m_{pl}}\right)^{2} \frac{4 m_{pl} \langle e^{2} \rangle^{2}}{\Sigma a^{2} \Omega},
\end{equation}
where $\Omega$, $a$ and $e$ are the orbital frequencies, semi-major axes and eccentricities of the individual planetesimals, \textbf{respectively}. In the Stokes regime, \textbf{where the mean-free path of the gas is much smaller than the solid particles, the gas can be treated as a fluid and} the drag timescale is given by \citet{adachi76} as
\begin{equation}\label{eq:ts_stokes}
    t_{s} = \frac{2 m_{pl}}{C_{D} \pi r_{pl}^{2} \rho_{g} v_{g}},
\end{equation}
where $C_{D}$ is a drag coefficient of order unity, $\rho_{g}$ is the local gas volume density and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. The local gas volume density is given by
\begin{equation}\label{eq:rho_gas}
	\rho_{g} = \frac{\Sigma_{g}}{\sqrt{2 \pi} h_{g}} \exp\left[ -z^{2} / \left( 2 h_{g}^{2} \right) \right],
\end{equation}
where $\Sigma_{g}$ is the gas surface density (taken to be
240 times the solid surface density \citep{hayashi81}, $h_{g} = c_{s} / \Omega$ is the local gas scale height and $z$ is the height above the disk midplane. The sound speed profile is given by $c_{s} = \sqrt{k_{B} T(r) / \left( \mu m_{H} \right)}$, where $k_{B}$ is Boltzmann's constant, $T(r) = T_{0} r^{-Q}$, $\mu = 2.34$ and $m_{h}$ is the mass of a hydrogen atom. For a protoplanetary disk around a typical M star, $T_{0} = 148$ K and $Q$ = 0.58 \citep{andrews05}.

@@

\begin{equation}\label{eq:v_gas}
	v_{g} = v_{k} \left[ 1 - \sqrt{ Q c_{s}^2 / v_{k}^2} \right],
\end{equation}

\noindent where $v_{k}$ is the local Keplerian velocity. As in section
\ref{sec:narrow}, the argument of perihelion $\omega$, longitude of
ascending node $\Omega$, and mean anomaly M for the planetesimals are drawn from a uniform distribution $\in [0, 2 \pi)$.

>>>>

where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g
cm$^{-2}$ is the surface density of the minimum-mass solar nebula
\citep{hayashi81} at 1 AU, $A$ is an enhancement factor and $\delta$ is the power law index.
In the first case (fdHi), we model a disk that follows a MMSN power law slope ($\delta$ = 1.5), with the overall normalization enhanced by a factor of 100. This choice of normalization for the solid surface density profile appears necessary in order to reproduce many observed short period terrestrial worlds in-situ \citep{hansen12}. \textbf{The most current planetesimal formation models all involve streaming instabilities triggered by solid material concentrating at preferential locations in the disk. This can occur via zonal flows \citep{johansen2009b, simon12}, vortices \citep{klahr03}, or through mechanisms that produce a pressure bump in the gas disk, such as an ionization \citep{lyra08} or condensation front \citep{brauer08b, drkazowska13}, or even the perturbations from an existing planet \citep{shibaike20}. \citet{drkazowska18} showed that evolution of the snow line boundary can cause planetesimal formation over a significant area of the disk, producing mass concentrations at least as large as the ones we use here. We argue that this is a particularly attractive mechanism for widespread planetesimal formation around low-mass stars, as their extreme pre-main sequence evolution is particularly conducive to significant movement of the condensation fronts \citep{baraffe15}.}

Additionally, we vary the power law index (fdHiShallow, fdHiSteep) and overall normalization (fdLo) of $\Sigma(r)$. \textbf{Although there is no way to reliably measure the uncertainty on the MMSN power law slope, \citet{chiang13} applied a similar analysis to a sample of Kepler multiplanet systems and found a variation of $\sim 0.2$. Sub-millimeter observations of the outer regions of cold protoplanetary disks find that $\delta$ can be as low as 0.5 \citep{mundy00, andrews09, andrews10}. Therefore, we vary $\delta$ by a value of $1.0$ relative to the MMSN value for the fdHiShallow and fdHiSteep simulations.}

@@

\begin{equation}\label{eq:vs_timescale}
    t_{vs}  = \frac{\left< e^2 \right>}{d \left< e^2 \right> / dt} \approx \frac{1}{40}\left(\frac{\Omega^{2} a^{3}}{2 G m_{pl}}\right)^{2} \frac{4 m_{pl} \langle e^{2} \rangle^{2}}{\Sigma a^{2} \Omega},
\end{equation}
where $\Omega$, $a$ and $e$ are the orbital frequencies, semi-major axes and eccentricities of the individual planetesimals, \textbf{respectively}. In the Stokes regime, \textbf{where the mean-free path of the gas is much smaller than the solid particles, the gas can be treated as a fluid and} the drag timescale is given by \citet{adachi76} as
\begin{equation}\label{eq:ts_stokes}
    t_{s} = \frac{2 m_{pl}}{C_{D} \pi r_{pl}^{2} \rho_{g} v_{g}},
\end{equation}
where $C_{D}$ is a drag coefficient of order unity, $\rho_{g}$ is the local gas volume density and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. The local gas volume density is given by
\begin{equation}\label{eq:rho_gas}
	\rho_{g} = \frac{\Sigma_{g}}{\sqrt{2 \pi} h_{g}} \exp\left[ -z^{2} / \left( 2 h_{g}^{2} \right) \right],
\end{equation}
where $\Sigma_{g}$ is the gas surface density (taken to be
240 times the solid surface density \citep{hayashi81}, $h_{g} = c_{s} / \Omega$ is the local gas scale height and $z$ is the height above the disk midplane. The sound speed profile is given by $c_{s} = \sqrt{k_{B} T(r) / \left( \mu m_{H} \right)}$, where $k_{B}$ is Boltzmann's constant, $T(r) = T_{0} \left( r / 1 \textrm{AU} \right)^{-q}$, $\mu = 2.34$ and $m_{h}$ is the mass of a hydrogen atom. For a protoplanetary disk around a typical M star, $T_{0} = 148$ K and $q$ = 0.58 \citep{andrews05}.

@@

\begin{equation}\label{eq:v_gas}
	v_{g} = v_{k} \left[ 1 - \sqrt{ q c_{s}^2 / v_{k}^2} \right],
\end{equation}

\noindent where $v_{k}$ is the local Keplerian velocity \textbf{(see eq. 4.30 of \citet{armitage20})}. As in section
\ref{sec:narrow}, the argument of perihelion $\omega$, longitude of
ascending node $\Omega$, and mean anomaly M for the planetesimals are drawn from a uniform distribution $\in [0, 2 \pi)$.
